---
title: Animating administrative data
author: ''
date: '2018-11-21'
slug: ageing-in-england
categories: []
tags: 
  - maps
  - sf
  - ageing
  - gganimate
subtitle: 'Looking at ageing in the English population'
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=14, fig.height=10,
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r, echo = FALSE, message = FALSE}

library(tidyverse)
library(sf)
library(viridis)
library(ggplot2)
library(gganimate)
library(readxl)
library(janitor)
library(knitr)
library(DT)

```


England as other developed countries is ageing. Is this process going to increase, stabilise, decrease ...? The Office of National Statistics regularly releases information on population trends. In this post I look at the [old-age dependency ratio](https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationestimates/articles/overviewoftheukpopulation/july2017) (OADR) which represents the people aged 65 and over for every 1,000 people aged 16 to 64 years (“traditional working age”).

The process is relatively nice and easy. It consists of three core steps: 

- Load the geographical data and select the information for England

- Arrange the data into a tidy format

- Plot the data in a faceted way

### Load the data 

First things first, let's get the geographical information (aka. the [shapefile](https://en.wikipedia.org/wiki/Shapefile)) on board. Before, this was a bit messy process but this is extremely easy with `sf::read_sf()`. The Office of National Statistics offers [geographical information](https://www.ons.gov.uk/methodology/geography/ukgeographies) at several levels for the whole United Kingdom. I am only interested in the information referred to the 326 districts in England (denominated with `E0...` )

```{r, echo  = TRUE, include= TRUE, message = FALSE}

england = sf::read_sf("./data/geography/england_lad_2017/Local_Authority_Districts_December_2017_Ultra_Generalised_Clipped_Boundaries_in_United_Kingdom_WGS84.shp") %>% 
  filter(str_detect(lad17cd, "^E0")) 


```


The data created is a tibble

```{r, echo  = FALSE, include= TRUE, message = FALSE}

head(england,5)

```



The second main source of information refers to the OADR. The data covers years 1996 to 2036^[Data for 2026 and 2036 are projections] in periods of 10 years. The names of the variables in the information of the ONS are not particularly clean so `janitor::clean_names()` does a splendid job here. 


```{r, echo  = TRUE, include= TRUE, message = FALSE}
ageing = read_excel("data/raw/ons_uk_ageing.xls", sheet = "Old Age Dependency Ratio" ) %>%
  clean_names() %>%
  filter(str_detect(area_code, "^E0")) 


```


### Tidy the data


The key is to present the data in long format having a `time` and a `dependency` variable for each decade and old age dependency ratio respectively. `tidyr::gather()` may be a good option here. 


```{r, echo  =TRUE, include= TRUE, message = FALSE}

ageing = ageing %>%
  gather(year, dependency, x1996:x2036, -area_code, -area_name) %>%
  mutate(year = gsub("x", "", year)) %>%
  arrange(area_code, year)

```

```{r, echo  = FALSE, include= TRUE, message = FALSE}
head(ageing,5)

```


Also, to fill the map I convert the OADR into a factor variable with `dplyr::ntile()` and link it to the data frame with geographical data (`england`). I create a variable `lab_dep` with the max and min value of each tile and that I will use as label for the represented variable. 


```{r, echo  = TRUE, include= TRUE, message = FALSE}
ageing = ageing %>%
  mutate(quintile_dep = ntile(dependency, 7), 
         lab_dep = as.factor(ordered(case_when(quintile_dep == 1 ~ "[81 - 218)", 
                                               quintile_dep == 2 ~ "[218 - 250)", 
                                               quintile_dep == 3 ~ "[250 - 283)",
                                               quintile_dep == 4 ~ "[283 - 329)",
                                               quintile_dep == 5 ~ "[329 - 392)", 
                                               quintile_dep == 6 ~ "[392 - 478)", 
                                               quintile_dep == 7 ~ "[478 - 928)")))) 


england_ext = full_join(england, ageing, by = c("lad17cd" = "area_code", "lad17nm" ="area_name") )

england_ext$quintile_dep = as.factor(england_ext$quintile_dep)

```

```{r, echo  = FALSE, include= TRUE, message = FALSE}
glimpse(england_ext)

```



### Plot the data


Let´s begin plotting the OADR for 2016. `sf::geom_sf()` seems to be a simpler solution to other alternatives such as `ggplot:geom_polygon()`.


```{r, echo  = TRUE, include= TRUE, message = FALSE}



 england_ext %>%
  filter(year == "2016") %>%
  mutate(lab_dep = gsub("[[:alpha:]]", "", lab_dep)) %>%
  rename(OADR = quintile_dep) %>%
  ggplot(aes(x = long, y = lat, fill = OADR)) + 
  geom_sf(colour = alpha("grey", 1 /3), size = 0.2) +
  coord_sf( datum = NA) +
  scale_fill_viridis(option = "viridis",
                     labels = c("[81 - 218)", "[218 - 250)", "[250 - 283)", "[283 - 329)", "[329 - 392)", "[392 - 478)","[478 - 928)"),
                     alpha = 0.85, 
                     discrete = T, 
                     direction = -1)

```

Not too bad.  Yet, what is the picture over time?.

```{r, echo  = TRUE, include= TRUE, message = FALSE}



plot_ageing = england_ext %>%
  mutate(lab_dep = gsub("[[:alpha:]]", "", lab_dep)) %>%
  rename(OADR = quintile_dep) %>%
  ggplot(., aes(x = long, y = lat, fill = OADR)) + 
  geom_sf(colour = alpha("grey", 1 /3), size = 0.2) +
  coord_sf( datum = NA) +
  scale_fill_viridis(option = "viridis",
                     labels = c("[81 - 218)", "[218 - 250)", "[250 - 283)", "[283 - 329)", "[329 - 392)", "[392 - 478)","[478 - 928)"),
                     alpha = 0.85, 
                     discrete = T, 
                     direction = -1, 
                     guide = guide_legend(
                       direction = "horizontal",
                       title.position = "top",
                       title.hjust =0.5)) +
  facet_wrap(~ year, ncol = 3)


```

```{r, echo  = FALSE, include= TRUE, message = FALSE}

plot_ageing

```

The key line is the `facet_wrap()` call which allows to facet `OADR` over `years`. A bit of fine tunning with some options in `theme()` give a cleaner plot.


```{r, echo  = TRUE, include= TRUE, message = FALSE}

plot_ageing +
  theme(axis.text = element_blank()
        ,axis.title = element_blank()
        ,axis.ticks = element_blank()
        ,axis.line=element_blank()
        ,panel.grid = element_blank()
        ,legend.text = element_text(size = 10)
        ,legend.key.width = unit(0.35,"cm")
        ,legend.key.height = unit(0.35,"cm")
        ,plot.title = element_text(size= 6)
        ,legend.position = "bottom"
        ,plot.caption = element_text()
        ,legend.background = element_blank()
        ,panel.background = element_blank()
        ,legend.spacing.x = unit(0.25, 'cm')
        ,strip.text.x = element_text(size = 12)) 


```

### Animate the plot 

We can combine the former plots by creating an animated map with `gganimate::animate()`. Here the main difference with respect to a simple plot in `ggplot()` is the incorporation of `transition_manual()` that specifies the transition in the map.

```{r, echo  = TRUE, include= TRUE, message = FALSE}

aniplot = england_ext %>%
  mutate(lab_dep = gsub("[[:alpha:]]", "", lab_dep)) %>%
  rename(OADR = quintile_dep) %>%
  ggplot(., aes(x = long, y = lat, fill = OADR)) + 
  geom_sf(colour = alpha("grey", 1 /3), size = 0.2) +
  coord_sf( datum = NA) +
  labs(title = "Old Age Dependency Ratio (OADR)",
       subtitle = 'Year: {current_frame}') +
  transition_manual(year) + 
  scale_fill_viridis(option = "viridis",
                     labels = c("[81 - 218)", "[218 - 250)", "[250 - 283)", "[283 - 329)", "[329 - 392)", "[392 - 478)","[478 - 928)"),
                     alpha = 0.85, 
                     discrete = T, 
                     direction = -1, 
                     guide = guide_legend(
                       direction = "horizontal",
                       title.position = "top",
                       title.hjust =0.5)) +
  theme(axis.text = element_blank()
        ,axis.title = element_blank()
        ,axis.ticks = element_blank()
        ,axis.line=element_blank()
        ,panel.grid = element_blank()
        ,legend.title = element_blank()
        ,legend.text = element_text(size = 14)
        ,legend.key.width = unit(0.35,"cm")
        ,legend.key.height = unit(0.5,"cm")
        ,plot.title = element_text(size= 22)
        ,plot.subtitle=element_text(size=18)
        ,legend.position = "bottom"
        ,plot.caption = element_text()
        ,legend.background = element_blank()
        ,panel.background = element_blank()
        ,legend.spacing.x = unit(0.25, 'cm'))

```


```{r, echo  = FALSE, include= TRUE, message = FALSE}

animate(aniplot, height = 1000, width = 1000)

```

The code and data to reproduce the figures are [available on Github](https://github.com/edugonzaloalmorox/dynamic_maps).
